---
title: "ST 558 Project 1"
author: "Marcus Lee"
date: "6/13/2021"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=TRUE, warning = FALSE, message=FALSE, include=FALSE}
library(ggplot2)
library(tidyverse)
library(jsonlite)
library(knitr)
library(httr)
```


# JSON Data

JSON, also known as [*Java Script Object Notation*](https://www.json.org/json-en.html) is a text based format for storing and transporting data. JSON is used widely across the internet and databases because of its text based format for storing data(Example: Getting data using API). In addition, JSON can represent 2D data, [hierarchical data](https://en.wikipedia.org/wiki/Hierarchical_database_model), and use key-value pairs.

Since JSON is a text based format, we need to load in R packages that can handle JSON data.

# Packages used for reading JSON data in R

Here are R packages that can be used to work with JSON data:

* `rjson`
* `RJSONIO`
* `jsonlite`

For this project, `rjson` and `jsonlite` is used.

Here are a couple resources for understanding the functionality of `rjson` and `jsonlite`:

1. [https://cran.r-project.org/web/packages/rjson/rjson.pdf](https://cran.r-project.org/web/packages/rjson/rjson.pdf)

2. [https://cran.r-project.org/web/packages/jsonlite/jsonlite.pdf](https://cran.r-project.org/web/packages/jsonlite/jsonlite.pdf)

# Contact the NHL Records API

To access the [NHL Records API](https://gitlab.com/dword4/nhlapi/-/blob/master/records-api.md), we need to construct a URL with the name of the table and attributes we want to pull from it.

In addition, I wrote 5 more functions that take different endpoints. In terms of preserving space in the README.md, you can find the 5 functions in the README.Rmd.

```{r}
base_url = "https://records.nhl.com/site/api"
```

```{r}
get_franchise = function(){
  full_url = paste0(base_url, "/franchise")
  franchise_text = content(GET(url=full_url), "text")
  franchise_json = fromJSON(franchise_text, flatten = T)
  return(franchise_json$data)
}
```

```{r, echo=FALSE, results='hide'}
get_franchise_team_totals = function(){
  full_url = paste0(base_url, "/franchise-team-totals")
  franchise_team_totals_text = content(GET(url=full_url), "text")
  franchise_team_totals_json = fromJSON(franchise_team_totals_text, flatten = T)
  return(franchise_team_totals_json$data)
}
```

```{r,echo=FALSE, results='hide'}
get_franchise_season_records = function(id){
  full_url = paste0(base_url, "/franchise-season-records?cayenneExp=franchiseId=",id)
  franchise_season_records_text = content(GET(url=full_url), "text")
  franchise_season_records_json = fromJSON(franchise_season_records_text, flatten = T)
  return(franchise_season_records_json$data)
}
```

```{r,echo=FALSE, results='hide'}
get_franchise_goalie_records = function(id){
  full_url = paste0(base_url, "/franchise-goalie-records?cayenneExp=franchiseId=",id)
  franchise_goalie_records_text = content(GET(url=full_url), "text")
  franchise_goalie_records_json = fromJSON(franchise_goalie_records_text, flatten = T)
  return(franchise_goalie_records_json$data)
}
```

```{r,echo=FALSE, results='hide'}
get_franchise_skater_records = function(id){
  full_url = paste0(base_url, "/franchise-skater-records?cayenneExp=franchiseId=",id)
  franchise_skater_records_text = content(GET(url=full_url), "text")
  franchise_skater_records_json = fromJSON(franchise_skater_records_text, flatten = T)
  return(franchise_skater_records_json$data)
}
```

```{r,echo=FALSE, results='hide'}
get_franchise_detail = function(id){
  full_url = paste0(base_url, "/franchise-detail?cayenneExp=mostRecentTeamId=",id)
  franchise_detail_text = content(GET(url=full_url), "text")
  franchise_detail_json = fromJSON(franchise_detail_text, flatten = T)
  return(franchise_detail_json$data)
}
```

Here is an use case of the `get_franchise` function working.

```{r, warning=FALSE, message =FALSE}
#This id represent the New Jersey Devils
id=23 
franchise = get_franchise()
franchise_team_total = get_franchise_team_totals()
franchise_season_records = get_franchise_season_records(id)
franchise_goalie_records = get_franchise_skater_records(id)
franchise_skater_records = get_franchise_skater_records(id)
franchise_detail = get_franchise_detail(id)
```

```{r}
kable(franchise)
```

# Contact the NHL Stats API

To access the [NHL Stats API](https://gitlab.com/dword4/nhlapi/-/blob/master/stats-api.md), we need to construct a URL with the name of the table and attributes we want to pull from it. Note that the url is different from the NHL Records API.

```{r}
base_stats_url = "https://statsapi.web.nhl.com/api/v1"
```

The function allows me to contact the [NHL Stats API](https://gitlab.com/dword4/nhlapi/-/blob/master/stats-api.md) for the ?expand=team.stats modifier. In this function, it should be able to take a single team or return data from all teams.

```{r}
get_team_stats = function(id){
  if (missing(id)==T){
    full_url = paste0(base_stats_url, "/teams?expand=team.stats=")
    team.stats_text = content(GET(url=full_url), "text")
    team.stats_json = fromJSON(team.stats_text, flatten = T)
    team.stats2none_json = team.stats_json[2] # We are ignoring the 1st list since it it useless
    return(team.stats2none_json)
  }
  else{
    full_url =  paste0(base_stats_url, "/teams/",id)
    team.stats_text = content(GET(url=full_url), "text")
    team.stats_json = fromJSON(team.stats_text, flatten = T)
    team.stats2id_json = team.stats_json[2] # We are ignoring the 1st list since it it useless
    return(team.stats2id_json)
  }
}
```

In this example, we set `id` to 2 and plug `id` into the `get_team_stats` function and assign it to `team.stats`. This should be to return data for a single team. However, in the code example, I did not specify an id meaning that `team.stat`s will get overwritten and will return data from all the teams instead.

```{r}
id=2
team.stats = get_team_stats(id)
# The code below here should also work if you wanted to print out the whole team.
team.stats = get_team_stats()
```

```{r}
kable(data.frame(team.stats))
```

## Wrapper function

The function below is a wrapper function that is a one-stop-shop for the user to access any of the API endpoints. In this case, this wrapper function should be able to get data from the NHL stats API function I wrote above and data from the NHL Records API function I wrote as well.

```{r}
get_wrapper <- function(api_url,id){
  if(api_url=="https://records.nhl.com/site/api"){
    base_url="https://records.nhl.com/site/api"
    if(missing(id)==T){
      franchise_wrapper = get_franchise()
      franchise_team_total_wrapper = get_franchise_team_totals()
      records_wrapper_no_id_list = list(franchise_wrapper,franchise_team_total_wrapper)
      return(records_wrapper_no_id_list)
    }
    else{
      franchise_season_records_wrapper = get_franchise_season_records(id)
      franchise_goalie_records_wrapper = get_franchise_skater_records(id)
      franchise_skater_records_wrapper = get_franchise_skater_records(id)
      franchise_detail_wrapper = get_franchise_detail(id)
      records_wrapper_list = list(franchise_season_records_wrapper,franchise_goalie_records_wrapper,franchise_skater_records_wrapper,franchise_detail_wrapper)
      return(records_wrapper_list)
    }
  }
  else{
    base_stats_url="https://statsapi.web.nhl.com/api/v1"
    if(missing(id)==T){
        team.stats = get_team_stats()
        return(team.stats)
    }
    else{
      team.stats = get_team_stats(id)
      return(team.stats)
    }
  }
}
```

Here is an use case of the `get_wrapper` function working. Note that I have tested my `get_wrapper` function with the `api_url` and `id` and all of the test cases work(The code for the test cases are in the README.Rmd file).

```{r,echo=FALSE, results='hide', message=FALSE, error=FALSE}
wrap = get_wrapper("https://records.nhl.com/site/api",2)
```

```{r,echo=FALSE, results='hide', message=FALSE, error=FALSE}
wrap = get_wrapper("https://records.nhl.com/site/api",)
```

```{r,echo=FALSE, results='hide', message=FALSE, error=FALSE}
wrap = get_wrapper("https://statsapi.web.nhl.com/api/v1",2)
```

```{r}
wrap = get_wrapper("https://statsapi.web.nhl.com/api/v1")
```

This will produce a table of the `team.stats` modifier from the NHL Stats API and return all the data from all teams. Note that I did not output the `wrap` because it is the same output as `kable(data.frame(team.stats))`.

```{r, echo=FALSE, results='hide', include=FALSE}
kable(data.frame(wrap))
```

# Exploratory Data Analysis

## Contigency Tables/ Numerical Summaries

Here is a table that shows a basic count of position codes for the New Jersey Devils by Active/non-active players based off of the `franchise_goalie_records` dataset.

```{r}
franchise_goalie_records$activePlayer = as.factor(franchise_goalie_records$activePlayer)
levels(franchise_goalie_records$activePlayer) = c("Not Active Player", "Active/Current Player")

franchise_goalie_records_table <- table(franchise_goalie_records$activePlayer, franchise_goalie_records$positionCode)

kable(franchise_goalie_records_table )
```

Here is a numerical summary of points made for the New Jersey Devils based off of position codes based off of the `franchise_goalie_records` data set. Based off of the numerical summary, the D position code does a way better job compared to the other position codes based off of total count. However, further analysis would need to be done to truly see if the D position code does a better job(Ex: Hypothesis test, etc.)

```{r}
numerical_table_summary <- franchise_goalie_records %>%
  group_by(positionCode)%>%
  summarize(
    Count = n(),
    Min = min(points, na.rm = T),
    Max = max(points, na.rm = T),
    Mean = mean(points, na.rm = T),
    Sd = sd(points, na.rm = T),
    percent_25 = quantile(points, probs=.25, na.rm = T),
    median = median(points, na.rm = T),
    percent_75 = quantile(points, probs=.75, na.rm = T)
  ) %>%
  rename(Code=positionCode, '25th percentile' = percent_25, '75th percentile' = percent_75)

kable(numerical_table_summary, caption = "Numerical Summaries of the New Jersey Devils Points by Position Code based off of the franchise goalie record data set.")
```

Here is another contingency table that shows a basic count of position codes for the New Jersey Devils by seasons based off of the `franchise_skater_records2` data set.

```{r}
franchise_skater_records2 <- franchise_skater_records
franchise_skater_records2$seasons = as.factor(franchise_skater_records2$seasons)

franchise_skater_records2_table <- table(franchise_skater_records2$positionCode, franchise_skater_records2$seasons)

kable(franchise_skater_records2_table)
```

Here is another numerical summary of `road_Win_Loss_Ratio` and `home_Win_Loss_Ratio` made for the New Jersey Devils based off of the teams based off of the `franchise_team_total` data set. 

Overall, when we take a look at the summary for the `home_Win_Loss_Ratio`, it appears that most of the teams performed way better in their home stadium as supposed to playing on their opponent's stadium.

```{r}
franchise_team_total2 <- franchise_team_total

franchise_team_total2 <- mutate(franchise_team_total2, road_Win_Loss_Ratio = roadWins/roadLosses)
franchise_team_total2 <- mutate(franchise_team_total2, home_Win_Loss_Ratio = homeWins/homeLosses)

get_summary <- function(z){
  if(z==1){
    franchise_team_total2 %>%
    group_by(teamName) %>%
    summarize(
      Min = min(road_Win_Loss_Ratio, na.rm = T),
      percent_25 = quantile(road_Win_Loss_Ratio, probs=.25, na.rm = T),
      Median = median(road_Win_Loss_Ratio, na.rm = T),
      Mean = mean(road_Win_Loss_Ratio, na.rm = T),
      percent_75 = quantile(road_Win_Loss_Ratio, probs=.75, na.rm = T),
      Max = max(road_Win_Loss_Ratio, na.rm = T),
      ) %>%
      rename("25th percentile"=percent_25, "75th percentile"=percent_75)
  }else if (z==2){
    franchise_team_total2 %>%
    group_by(teamName) %>%
    summarize(
      Min = min(home_Win_Loss_Ratio, na.rm = T),
      percent_25 = quantile(home_Win_Loss_Ratio, probs=.25, na.rm = T),
      Median = median(home_Win_Loss_Ratio, na.rm = T),
      Mean = mean(home_Win_Loss_Ratio, na.rm = T),
      percent_75 = quantile(home_Win_Loss_Ratio, probs=.75, na.rm = T),
      Max = max(home_Win_Loss_Ratio, na.rm = T),
      ) %>%
      rename("25th percentile"=percent_25, "75th percentile"=percent_75)
  }else{
    print("Wrong input!!!")
  }
}
franchise_team_total2_numerical_road_Win_Loss_Ratio_summary <-get_summary(1)
franchise_team_total2_numerical_home_Win_Loss_Ratio_summary <- get_summary(2)

head(franchise_team_total2_numerical_road_Win_Loss_Ratio_summary)
head(franchise_team_total2_numerical_home_Win_Loss_Ratio_summary)
```


## Graphical Summaries

Continuing with the `road_Win_Loss_Ratio` & `home_Win_Loss_Ratio` summary, here are a couple plots that can help us have a better understanding of the `road_Win_Loss_Ratio` & `home_Win_Loss_Ratio`.

Here is a box plot that shows the Ratio of Road Win to Losses. Note, that the reason we did not make box plots for each team instead because if we were to make one, the scale would be off and it would be hard to tell since there are a lot of teams represented(There are not a lot of colors).

```{r}
ggplot(franchise_team_total2, aes(road_Win_Loss_Ratio, y=""))+geom_boxplot(fill="blue") + ggtitle("Ratio of Road Win to Losses Boxplot") + xlab("Ratio of Road Win to Losses") 
```

Here is another box plot, but with the Ratio of Home Win to Losses

```{r, warning=FALSE}
ggplot(franchise_team_total2, aes(home_Win_Loss_Ratio, y=""))+geom_boxplot(fill="green") + ggtitle("Ratio of Home Win to Losses Boxplot") + xlab("Ratio of Home Win to Losses") 
```

By comparing the two box plots, we can see that teams won a lot at their home stadium than playing at their opponent team's stadium.

Since box plots have limits on showing certain statistics, we might consider looking at histograms of the ratios.

However, one should be careful because changing the binwidth can change the appearance of the histogram leading to wrong conclusions. Other than that, it appears most of the ratios of road win to road losses tend to be around 0.70 to 0.90. 

```{r}
ggplot(franchise_team_total2, aes(x=road_Win_Loss_Ratio))+geom_histogram(aes(y=..density..),color="darkblue", fill="lightblue", binwidth = 0.10) + geom_density(kernel="gaussian", color="red")+ ggtitle("Histogram of Road Win to Road Losses")+ylab("Density")
```

Since box plots and histograms have limits on showing certain statistics, we might consider looking at bar plots of the `roadWins` and `roadLosses` by `id` on the `franchise_team_total2` data set.

By looking and comparing the two bar plots, we should realize that it is hard to come up with a conclusion with the bar plots because some ids belong to certain teams. Not only just that, but it is hard whether or not the teams made an improvement on increasing their `roadWins` or decreasing their `roadLosses` since the data set does not have a time component. Moreover, the bar plots provided down below do not give as much statistical information compared to a box plots and histograms.

```{r}
ggplot(franchise_team_total2, aes(x=id, y=roadWins))+geom_bar(stat="identity", fill="steelblue",width=0.9)+ggtitle("Bar plot of id roadWins")
```

```{r}
ggplot(franchise_team_total2, aes(x=id, y=roadLosses))+geom_bar(stat="identity", fill="lightblue",width=0.9)+ggtitle("Bar plot of id roadLosses")
```

Now, let us examine to see if there is some type of relationship between Road Wins and Home Wins since we took a look at `road_Win_Loss_Ratio` & `home_Win_Loss_Ratio plots`.

Based off of this scatterplot, it shows that a lot of teams(regardless of `gameTypeId`) got more home wins as supposed to road wins because most of the points are above the line. However, although it might seem like playing in a home stadium gives the team more advantage as supposed to playing in an opponent's stadium, we need to do further Exploratory Data Analysis(Transforming data, statistical analysis, etc.).

```{r}
ggplot(franchise_team_total2, aes(x=roadWins,y=homeWins, color=gameTypeId))+geom_point()+geom_abline()+labs(x="Road Wins", y="Home Wins", title="Scatterplot of Road Wins vs Home Wins")
```





































































